---
title: "Proposal_Code"
output: pdf_document
---

1. Clean Data
```{r}
# Read the dataset
data <- read.csv("nba_final.csv")

# Subset the data
data <- data[, c("PTS", "Role", "Age", "FGA", "FT.", 
                 "X2P.", "X3P.", "TOV", "PF")]

# Remove rows with missing values (NA)
data <- na.omit(data)

# Create the Age_Category column
data$Age_Category <- ifelse(data$Age < 27, "Pre-Peak Age", 
                            ifelse(data$Age >= 27 & data$Age <= 29,
                                   "Peak Age", "Post-Peak Age"))

# It seems that R treat Age_Category and Role as character, convert character to factor
data$Role <- factor(data$Role)
data$Age_Category <- factor(data$Age_Category)

# B.3 Table of numerical summaries
summary(data)
```
1. Extra: Check Skewed Variable
```{r}
# Create a 3x3 grid of plots
par(mfrow = c(3, 3))

# Create frequency vs. variable graphs for each variable
variables <- c("PTS", "Role", "Age", "FGA", "FT.", "X2P.", "X3P.", "TOV", "PF")

for (variable in variables) {
  if (is.numeric(data[, variable])) {
    # Histogram for numerical variables
    hist(data[, variable], main = paste("Frequency vs.", variable), 
         xlab = variable, ylab = "Frequency")
  } else {
    # Barplot for categorical variables
    barplot(table(data[, variable]), main = paste("Frequency vs.", variable), 
            xlab = variable, ylab = "Frequency")
  }
}

# Reset par() settings to default
par(mfrow = c(1, 1))

```

2. Fit Multiple Linear Regression Model:
```{r}
# Fit the multiple linear regression model
model <- lm(PTS ~ Role + Age_Category + FGA + FT. + X2P. + X3P. 
            + TOV + PF + FGA:Role, data = data)

variable_descriptions <- c("Intercept", 
                           "Role: Front", 
                           "Age at Pre-Peak",
                           "Age at Post-Peak",
                           "Field Goal Attempts Per Game", 
                           "Free throw percentage", 
                           "Field Goal Percentage on 2-Pt Field Goal Attempts Per Game", 
                           "Field Goal Percentage on 3-Pt Field Goal Attempts Per Game", 
                           "Turnovers Per Game", 
                           "Personal Fouls Per Game", 
                           "Interaction btw FGA and Front Role")
# Print the model coefficients
model
# Print the coefficients with labels
coefficients_with_labels <- cbind(Variable = variable_descriptions, 
                                  Coefficient = coef(model))
print(coefficients_with_labels)
```
3. Check Assumptions of MLR Model:
a. Residuals versus each predictor (scatterplot for numerical / boxplot for categorical):
```{r}
# Calculate residuals
residuals <- residuals(model)

# Create a layout for separate plots
layout(matrix(1:6, nrow = 2))  # Create a 2x3 grid of plots

# Create scatterplots of residuals versus each predictor
plot(residuals ~ data$FGA, main = "Field Goal Attempts \n Per Game", 
     xlab = "FGA", ylab = "Residuals")
plot(residuals ~ data$FT., main = "Free throw percentage", 
     xlab = "FT.", ylab = "Residuals")
plot(residuals ~ data$X2P., main = "Percentage on 2Pt Goal \n Per Game", 
     xlab = "X2P.", ylab = "Residuals")
plot(residuals ~ data$X3P., main = "Percentage on 3Pt Goal \n Per Game", 
     xlab = "X3P.", ylab = "Residuals")
plot(residuals ~ data$TOV, main = "Turnovers Per Game", 
     xlab = "TOV", ylab = "Residuals")
plot(residuals ~ data$PF, main = "Personal Fouls Per Game", 
     xlab = "Role", ylab = "PF")

# Create scatterplots of residuals versus each predictor
boxplot(residuals ~ data$Role, main="Residual vs Role", 
        xlab="Role", ylab="Residual")
boxplot(residuals ~ data$Age_Category, main="Residual vs Age Category", 
        xlab="Age Category", ylab="Residual")
```
b. Residual versus fitted values (scatterplot):
```{r}
# Create a scatterplot of residuals versus fitted values
plot(fitted(model), residuals,  main = "Residuals vs. Fitted", 
     xlab = "Fitted Values", ylab = "Residuals")
abline(h = 0, col = "red")  # Add a horizontal line at y = 0 for reference

```
c. Normal Quantile-Quantile (QQ) Plot:
```{r}
# Create a normal QQ plot of residuals
qqnorm(residuals)
qqline(residuals, col = "red")  # Add a reference line
```
4. Check additional Condition of MLR Model:
a. Conditional Mean Response Condition (Linearity of the Mean Response):
```{r}
# Create a scatterplot of response vs. fitted values
plot(x = fitted(model), y = data$PTS, main = "Response vs. Fitted", 
     xlab = "Fitted Values", ylab = "Response")
abline(a = 0, b = 1, col = "red") # diagonal line for reference
```
b. Conditional Mean Predictor Condition (Linearity among Predictors):
```{r}
# Create pairwise scatterplots of predictors
pairs(data[, c("FGA", "FT.", "X2P.", "X3P.", "TOV", "PF", "Age_Category", "Role")])
```

Export cleaned dataset
```{r}
write.csv(data, file = "cleaned_data.csv", row.names = FALSE)
```

